---
title: "Analyzing GDP for final project in Software Carpentry Workshop Oct 2015"
author: "Anna Dornhaus"
date: "10/4/2015"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE}
library("dplyr")

```

We can see that the GDP per Person has been clearly increasing in the countries Germany, Mexico, and Mongolia.

```{r read-in-data, echo=FALSE}
gapdata <- read.table("output/combined_gapMinder.tsv", sep = "\t", header = TRUE)
```

```{r calculations, echo=FALSE}
GermanyData <- subset(gapdata, country=="Germany")
MexicoData <- subset(gapdata, country=="Mexico")
MongoliaData <- subset(gapdata, country=="Mongolia")
```

```{r plotting, echo=FALSE}
plot(gdpPercap ~ year, 
     data=GermanyData,
     xlab = list("Year"),
     ylab = list("GDP per Person"),
     main = "GDP Change for Germany"
     )
plot(gdpPercap ~ year, 
     data=MexicoData,
     xlab = list("Year"),
     ylab = list("GDP per Person"),
     main = "GDP Change for Mexico"
     )
plot(gdpPercap ~ year, 
     data=MongoliaData,
     xlab = list("Year"),
     ylab = list("GDP per Person"),
     main = "GDP Change for Mongolia"
     )

```

Life expectancy by continent:
```{r statistics, echo=FALSE}
AvgLifespan <- aggregate(gapdata$lifeExp, list(gapdata$continent), mean)
MinLifespan <- aggregate(gapdata$lifeExp, list(gapdata$continent), min)
MaxLifespan <- aggregate(gapdata$lifeExp, list(gapdata$continent), max)

colnames(AvgLifespan) <- c("Continent", "Average Lifespan")
colnames(MinLifespan) <- c("Continent", "Minimum Lifespan")
colnames(MaxLifespan) <- c("Continent", "Maximum Lifespan")
```

We clearly see that maximum life expectancy is greater than the mean.
```{r lifeExpoutput, echo=FALSE}
AvgLifespan
MinLifespan
MaxLifespan

```

Now we'll work on histograms.
This shows that a life expectancy of around 70 is most frequent across countries, but a fair number have life expectancies below 50 in some years.
```{r histograms, echo=FALSE}
hist(gapdata$lifeExp, main = "Distribution across countries of life expectancy")
hist(gapdata$lifeExp, breaks = 40, main = "Distribution across countries of life expectancy")

data2007 <- subset(gapdata, year==2007)
hist(data2007$lifeExp, main = "Distribution across countries of life expectancy for 2007 (no duplicate entries for countries)")

```

Here's the end of our analysis output.

